\documentclass[12pt,a4paper,oneside]{report}

%----------------------------------------------------------------------------------------
%   PACKAGES & BASIC SETUP
%----------------------------------------------------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{microtype}
\usepackage{setspace}
\onehalfspacing
\usepackage{listings}

\usepackage[margin=1in]{geometry}
\setlength{\parindent}{0em}
\setlength{\parskip}{0.5em}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{longtable}

%----------------------------------------------------------------------------------------
%   HARVARD (AUTHOR-DATE) REFERENCING
%----------------------------------------------------------------------------------------
\usepackage[round,authoryear]{natbib}
% 'round' gives (Author, Year); 'authoryear' ensures the author-date format.
\bibliographystyle{agsm}  % Use Harvard-like style for the bibliography.

% Custom macros for consistent referencing
\newcommand{\citeauthorstyle}[1]{\citeauthor{#1} (\citeyear{#1})}  % Author (Year)
\newcommand{\citeparensstyle}[1]{(\citeauthor{#1}, \citeyear{#1})} % (Author, Year)

%----------------------------------------------------------------------------------------
%   HYPERLINKS (Load Last)
%----------------------------------------------------------------------------------------
\usepackage[colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue]{hyperref}

%----------------------------------------------------------------------------------------
%   GLOSSARIES PACKAGE FOR ABBREVIATIONS
%----------------------------------------------------------------------------------------
\usepackage[acronym]{glossaries}

% Define abbreviations
\newacronym{twod}{2D}{two dimensional}

%----------------------------------------------------------------------------------------
%   TITLE AND AUTHOR COMMANDS
%----------------------------------------------------------------------------------------
\newcommand{\reporttitle}{Solar Flare Prediction with Multimodal Transformer Architectures: A Novel Approach to Interpretable Space Weather Forecasting}
\newcommand{\reportauthor}{Antanas Žilinskas}
\newcommand{\reporttype}{Design Engineering Master’s Project}

% Adjust ToC Depth if desired
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{2}

\begin{document}

%----------------------------------------------------------------------------------------
%   TITLE PAGE
%----------------------------------------------------------------------------------------
%TC:ignore
\input{titlepage}
%TC:endignore

\clearpage
\pagenumbering{roman}

%----------------------------------------------------------------------------------------
%   DECLARATION OF ORIGINALITY & COPYRIGHT DECLARATION
%----------------------------------------------------------------------------------------
%TC:ignore
\clearpage
\section*{Declaration of Originality}

\noindent
I hereby declare that the work presented in this thesis is my own, except where explicitly stated otherwise. To the best of my knowledge, it is original, and any collaborative contributions or referenced ideas have been duly acknowledged. Large Language Models (e.g., Claude and ChatGPT) were utilized to aid with project development, enhance the quality of writing, and as tools during research.

\vspace{2cm} % Space between the two declarations

\section*{Copyright Declaration}

\noindent
The copyright of this thesis rests with the author and is made available under a Creative Commons Attribution Non-Commercial No Derivatives licence. Researchers are free to copy, distribute, or transmit the thesis on the condition that they attribute it, that they do not use it for commercial purposes, and that they do not alter, transform, or build upon it. For any reuse or redistribution, researchers must make clear to others the licence terms of this work.
%TC:endignore

%----------------------------------------------------------------------------------------
%   ABSTRACT
%----------------------------------------------------------------------------------------
%TC:ignore
\clearpage
\section*{Abstract}
Solar flare forecasting is a critical challenge at the intersection of space weather prediction and advanced computational modeling. Current state-of-the-art approaches, including statistical methods, traditional machine learning models \cite{RefWorks:RefID:12-zheng2023comparative}, and deep learning architectures based on convolutional and recurrent networks \cite{RefWorks:RefID:10-jiao2020solar}, provide incremental improvements but often struggle with the intrinsic complexity, multimodality, and nonlinearity of solar data. These shortcomings restrict both predictive accuracy and lead times, especially for the rare but highly impactful X-class flares.

The following research propose a plan for a novel transformer-based methodology inspired by Large Language Models (LLMs) \cite{RefWorks:RefID:3-vaswani2023provided,RefWorks:RefID:2-abduallah2023operational} to address these limitations. By integrating spatial magnetogram imagery, time-series flux measurements, and related solar observations into a unified, self-attention-driven framework \cite{RefWorks:RefID:35-licllmate:}, approach aims to set new performance benchmarks. Research will employ self-supervised pretraining to learn generalizable representations of solar phenomena \cite{RefWorks:RefID:29-hoffmanntraining}, thereby overcoming data scarcity and improving model robustness. The resulting architecture seeks not only to enhance predictive skill but also to offer interpretability through attention mechanisms \cite{RefWorks:RefID:17-devlinbert:}, granting insights into which features and temporal intervals are most indicative of impending events.

The anticipated contributions include more accurate, early detection of major solar flares, improved understanding of their underlying drivers, and guidance for space weather stakeholders to better protect infrastructure and human activity dependent on reliable technological systems. Ultimately, this work aspires to establish a new standard for solar flare predictions using transformer-based architecture.
\noindent
%TC:endignore

%----------------------------------------------------------------------------------------
%   TABLE OF CONTENTS
%----------------------------------------------------------------------------------------
%TC:ignore
\clearpage
\tableofcontents
%TC:endignore

%----------------------------------------------------------------------------------------
%   ABBREVIATIONS, ACRONYMS, AND INITIALISMS
%----------------------------------------------------------------------------------------
% Uncomment if needed and exclude from word count
% %TC:ignore
% \clearpage
% \section*{Abbreviations, Acronyms, and Initialisms}
% \printglossary[type=\acronymtype, title={}]
% %TC:endignore

%----------------------------------------------------------------------------------------
%   FIGURES AND TABLES (OPTIONAL)
%----------------------------------------------------------------------------------------
% Uncomment if you have figures or tables and exclude from word count
% %TC:ignore
% \clearpage
% \listoffigures
% \clearpage
% \listoftables
% \clearpage
% %TC:endignore

\pagenumbering{arabic}
\setcounter{page}{1}

%----------------------------------------------------------------------------------------
%   CHAPTERS
%----------------------------------------------------------------------------------------
\chapter{Introduction}
\label{chap:intro}
\input{Chapters/1-introduction}

\chapter{Literature Review and Contextual Investigation}
\label{chap:literature-review-and-contextual-investigation}
\input{Chapters/2-literature-review-and-contextual-investigation}

\chapter{Project Objectives and scope}
\label{chap:project-objectives-and-scope}
\input{Chapters/3-project-objectives-and-scope}

\chapter{Methodological Approach}
\label{chap:methodological-approach}
\input{Chapters/4-methodological-approach}

\chapter{Project Management and Timeline}
\label{chap:project-management-and-timeline}
\input{Chapters/5-project-management-and-timeline}

\chapter{Expected Contributions and Impact}
\label{chap:expected-contributions-and-impact}
\input{Chapters/6-expected-contributions-and-impact}

\chapter{Conclusion and Next Steps}
\label{chap:conclusion-and-next-steps}
\input{Chapters/7-conclusion-and-next-steps}

%----------------------------------------------------------------------------------------
%   REFERENCES
%----------------------------------------------------------------------------------------
%TC:ignore
\clearpage
\bibliographystyle{agsm}  % Harvard-like style
\bibliography{References}
%TC:endignore

\end{document}
